---
title: "PSTAT 10 Review Sheet"
author: "Tianle Yu"
date: "2022-11-16"
output:
  html_document:
    toc: yes
    theme: readable
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data in R

## Datatype
\
* integer\
* double\
* logical\
* char\

```{r}
a <- "Hi"
typeof(a)
a <- 1
typeof(a)
a <- 1L
typeof(a)
```

## Data Structure
\
 - Vector\
 - Matrix\
 - Array\
 - Dataframe\
 - List\
 - Factor\

### Vector

```{r}
# Vector:
v <- 1:6
typeof(v)
v <- c(1,2,3,4,5,6)
typeof(v)
is.vector(v)
class(v)
```

### Matrix

```{r}
# Matrix
# Create 4x3 matrix
m <- matrix(1:5,nrow=4,ncol=3,byrow=F)
m
(1:10)+(2:5)
dim(m)
m[1:3,2]
rownames(m) <- c("A","B","C","D")
m
m["A",]
m[c("A","B"),]
m[c(1,2),]
```

### Array

```{r}
# Array
a <- array(1:120,dim=c(5,5,3))
a[2,2,2]
dimnames(a) <- list(c("A","B","C","D","E"),c("A","B","C","D","E"),c("A","B","C"))
# a
a[c(3,5),2,]
```

### Dataframe

```{r}
# Dataframe
d <- data.frame(1:3,c("A","B","C"),factor(c("D","D","A")),c("F","G",NA))
#d
d[1]
d[[1]]
colnames(d) <- c("F","G","H")
d
d[c(T,F,T),]
d[d$F==2,]
```

### List

```{r}
# List
l <- list(1:5,c("A","B"),factor(c("A","A","B","C")))
l
l[[1]]
l[[2]][2]
```

### Factor

```{r}
# Factor
# Unordered
factor(c("A+","A","A"))

# Ordered: order by ascii order
factor(c("A+","A","A-"),ordered = T)

# Ordered: order by assign order
factor(c("A+","A","A","A-"),ordered = T,levels = c("A-","A","A+"))

```

# Control Structure

## Variable Name
```{r, error=T}
a <- 1
b <- "123"
ABC <- 2
abc <- 234
sdf2 <- 23234
2432ds <- 342
342 <- 344534
fhdus_ <- 3242
_fdusif <- 344
fhus_dsaf <- 2453
for <- 3423
```

## Logical
```{r}
if(1==2){
  print("NMSL")
}else{
  print("NMHL")
}
```

## Loop
### While
```{r}
x <- 10
while(x>0){
  print(x)
  x <- x-1
}
```
### For
```{r}
fdsfds = c("a","b","c")
for(i in fdsfds){
  print(i)
}
```
## Function
```{r}
absoluteVal <- function(parameter){
  if(!is.numeric(parameter))
    return(FALSE)
  
  if(parameter<0)
    return(-parameter)
  else
    return(parameter)
  
}
myfunc <- function(input){
  if(input<10)
    return(input)
  else if(input<20){
    input <- input + 10
    return(input)
  }
  else{
    input <- input -10
    return(input)
  }
}

absoluteVal("Dgfdsg")
absoluteVal(2)
absoluteVal(-3)
```

# Plot & Graph
## Import Dataframe
### From csv
```{r}
df <- read.csv("./sleep.csv")
```
### From database

## Graphs
### Plot
Graph of points, relation between X and Y
```{r}
plot(df$age,df$sleep,xlab = "Age",ylab="Hours of sleep",main="Age vs. hours of sleep",pch=1)
```

### Histogram
Distribution of certain continuous (numeric) column
```{r}
hist(df$age)
```

### Barplot
Discrete data / Categorical data distribution / count
```{r}
barplot(table(df$standing),main = "123",col=c("red","green"))
```

### Boxplot
```{r}
summary(df$age)
```

```{r}
boxplot(df$age)
```
```{r}
boxplot(df$sleep~df$gender,col=c("red","green","purple"))
```

# Probability
## Range\
$\mathbb{P}(x) \in [0,1]$\

## Calculation
$\mathbb{P}(x \cap y)$\
$\mathbb{P}(x \cup y) = \mathbb{P}(x) + \mathbb{P}(y) - \mathbb{P}(x \cap y)$\


$\mathbb{P}(x|y) = \frac{\mathbb{P}(x \cap y)}{\mathbb{P}(y)}$

## Property
Muturally Exclusive\
$\mathbb{P}(x \cap y) = 0$\

Independent\
$\mathbb{P}(x \cap y) = \mathbb{P}(x) \cdot \mathbb{P}(y)$\

# Distribution
## Functions
 - P.M.F.  Probability mass function\
 - C.D.F.  Cumulative distribution function\
 - P.D.F.  Probability Density Function\
 
## R Functions
 - dxxx $\rightarrow$ P.M.F or P.D.F\
 - pxxx $\rightarrow$ C.D.F
 - rxxx $\rightarrow$ Run experiment
 - qxxx $\rightarrow$ Inverse of pxxx
 
## Uniform Distribution
 - Constant probability\
 - Continous\
A bus arrives at a stop every 10 minutes. A student is equally
likely to arrive at the stop at any time. How long will the student
have to wait?\

$\mathbb{P}(X \leq x)$ $\rightarrow$ C.D.F(x) $\rightarrow$ punif(x,min,max)\
Prob of waiting for 3 mins or less\
$\mathbb{P}(X \leq 3)$\
```{r}
punif(3,0,10)
```

Prob of waiting for 3-5 mins\
$\mathbb{P}(3 \leq X \leq 5) = \mathbb{P}(X \leq 5) - \mathbb{P}(X \leq 3)$\
```{r}
punif(5,0,10) - punif(3,0,10)
```

```{r}
punif(0:10,0,10)
```
## Binomial Distribution
 - Constant number of trials\
 - Two outcomes for every trial: Success or Fail\
 - Constant probability between trials\
 - Trails need to independent\

$\mathbb{P}(X=x)$ $\rightarrow$ P.M.F(x) $\rightarrow$ dbinom(x,trials,probability of success)\
$\mathbb{P}(X \leq x)$ $\rightarrow$ C.D.F(x) $\rightarrow$ pbinom(x,trials,probability of success)\
```{r}
plot(dbinom(0:10,10,0.5))
```

Toss a coin 10 times, getting 5 heads
```{r}
dbinom(5,10,0.5)
```

Toss a coin 10 times, getting less than 5 heads
$\mathbb{P}(X \leq 5)$ $\rightarrow$ C.D.F(5) $\rightarrow$ pbinom(5,n,p)\
```{r}
sum(dbinom(0:4,10,0.5))
pbinom(4,10,0.5)
cumsum(dbinom(0:10,10,0.5))[5]
```

$\mathbb{P}(X \leq 4.1)$ $\rightarrow$ C.D.F(4.1) $\rightarrow$ pbinom(4.1,n,p)\
```{r}
pbinom(4.1,10,0.5)
plot(pbinom(0:10,10,0.5),type="s")
```
Probability of getting 3-7 heads (inclusive)
```{r}
pbinom(7,10,0.5)-pbinom(2,10,0.5)
pbinom(7,10,0.5)-pbinom(3,10,0.5) + dbinom(3,10,0.5)
```

## Normal Distribution

# SQL


